---
title: "433HW2"
author: "Yifan Zhang"
output: github_document
---


## Load necessary libraries and look at the data


```{r}
library(dplyr)
library(nycflights13)
library(ggplot2)

flights
?flights
flights[,5:10]
colnames(flights)
str(flights)
```

## Problem 1

How many flights have a missing dep_time? What other variables are missing? What might these rows represent?

```{r}
f <- flights

#flights missing a dep_time
m <- f[is.na(f$dep_time), ]
nrow(m)

#other missing variables
head(m)
summary(m)
# From the summary, we can see in these rows variables dep_delay, arr_time, arr_delay and air_time also miss.

#Representation
#I think these rows represent those cancelled flights. Because they are cancelled, they don't have scheduled and actual departure and arrival times as well as amount of time spent in air.
```


## Problem 2

Currently dep_time and sched_dep_time are convenient to look at, but hard to compute with because they’re not really continuous numbers. Convert them to a more convenient representation of number of minutes since midnight.

```{r}
minutes <- function(x){
  for (i in 1: length(x)) {
    if(!is.na(x[i])){
      if(nchar(x[i]) == 1){
        x[i] <- x[i]
      }else if(nchar(x[i]) == 2){
        x[i] <- x[i]
      }else if(nchar(x[i]) == 3){
        x[i] <- as.numeric(substr(x[i], 1, 1)) * 60 + as.numeric(substr(x[i], 2, 3))
      }else{
        x[i] <- as.numeric(substr(x[i], 1, 2)) * 60 + as.numeric(substr(x[i], 3, 4))
      }
    }
  }
  return(x)
}

f$dep_time <- minutes(f$dep_time)
f$sched_dep_time <- minutes(f$sched_dep_time)
head(f)
```

## Problem 3

Look at the number of cancelled flights per day. Is there a pattern? Is the proportion of cancelled flights related to the average delay? Use multiple dyplr operations, all on one line, concluding with ggplot(aes(x= ,y=)) + geom_point()

```{r}
f %>% 
  group_by(month, day) %>%
  summarize(averageDelay = mean(dep_delay, na.rm = TRUE),
            proportionCancelled = sum(is.na(dep_time) / n())) %>%
  ggplot(aes(x = averageDelay, y = proportionCancelled)) +
  geom_point(size = 1, col = "#2c3269") +
  geom_smooth(se = FALSE, col = "#9b161b") +
  labs(x = "Average Delay", 
       y = "Proportion Cancelled")

#From the graph, we see that the proportion of cancelled flights is positively related to the average delay.
```
